### 2022-10-26 characterisation/Code/Comparing_symptoms.Rmd
update

---
title: "Exploring_symptoms_across_cohorts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message=FALSE, 
                      comment = FALSE,
                      cache.lazy=FALSE)
options(scipen=999)
```

```{r include=FALSE}
# packages
library(SqlRender)
library(DBI)
library(DatabaseConnector)
library(here)
library(dplyr)
library(tidyr)
library(CDMConnector)
library(stringr)
library(lubridate)
library(gtsummary)

# Connection details ----
server    <- Sys.getenv("SERVER_JUN22")
server_dbi<- Sys.getenv("SERVER_DBI_JUN22")
user      <- Sys.getenv("DB_USER_JUN22")
password  <- Sys.getenv("DB_PASSWORD_JUN22")
port      <- Sys.getenv("DB_PORT_JUN22") 
host      <- Sys.getenv("DB_HOST_JUN22")

db <- dbConnect(RPostgres::Postgres(), 
                dbname = server_dbi,
                port = port,
                host = host, 
                user = user, 
                password = password)

targetDialect              <- "postgresql"
cdm_database_schema        <- "omop21t4_cmbd" 
vocabulary_database_schema <-"omop21t4_cmbd" 
write_schema               <-"results21t4_cmbd"

# Generate cdm object to access the tables
cdm <- cdm_from_con(db, 
             cdm_schema = cdm_database_schema,
             write_schema = write_schema,
             cohort_tables = c("er_long_covid_final_cohorts",
                               "er_long_covid_all_symptoms_cohorts"))

```

```{r include=FALSE}
generate_table <- function(cohort_id, window, name){
data <- cdm$er_long_covid_all_symptoms_cohorts %>%
  filter(main_cohort_id == cohort_id & window_days == window) %>%
  collect()
summary <-  data%>% 
  mutate(all_symp = as.factor(all_symp)) %>%
  select(-subject_id, -window_days, - main_cohort_id, - cohort_start_date ) %>%
  mutate(gen = paste0(name)) %>%
   tbl_summary(
       by = "gen",
  digits = all_categorical()~ c(0,1),
  missing = "ifany",
  missing_text= "Missing",
  percent = "col"
           ) %>% 
  bold_labels()
  summary
}
```

```{r include=FALSE}
t.covid.infection <- generate_table(cohort_id= "22",
                                    window="90",
                                    name = "New COVID-19 infection")

t.covid.reinfections <- generate_table(cohort_id= "71",
                                    window="90",
                                    name = "Reinfections")

t.covid.confirmed <- generate_table(cohort_id= "30",
                                    window="90",
                                    name = "Confirmed infections")

t.tested.negative.earliest <- generate_table(cohort_id = "29",
                                    window="90",
                                    name = "Tested negative, earliest")

t.pcr.negative.earliest <- generate_table(cohort_id = "24",
                                    window="90",
                                    name = "PCR negative, earliest")

```

```{r}

table  <- tbl_merge(tbls = list(t.covid.infection,      
                                t.covid.reinfections,
                                t.tested.negative.earliest,
                                t.pcr.negative.earliest
                                 ), 
                     tab_spanner = c("", "", "", ""))

table

```
```{r}
## explorar numero de sintomas en tested negative y covid infection
# plot_grid(
#   covid_infection_numb_symp %>% filter(window==90) %>% 
#     ggplot(aes(all_symp)) +geom_bar() + theme_minimal() + ggtitle("Covid-19")  ,
# tested_negative_all_numb_symp %>% filter(window==90) %>%
#   ggplot(aes(all_symp)) +geom_bar() + theme_minimal() + ggtitle("Tested negative")
# )

```