### 2022-11-17 characterisation/Code/Comparing_symptoms_matched.Rmd
update

---
title: "Descriptive characteristics of Long COVID cohrots, matched"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning=FALSE, 
                      message=FALSE, 
                      comment = FALSE,
                      cache = FALSE,
                      cache.lazy=FALSE)
options(scipen=999)

```

```{r include=FALSE}
# packages
library(here)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(gtsummary)
library(tableone)
library(kableExtra)
library(ggplot2)
library(forcats)
library(tsibble)
library(survey)

# get data for table 1
# get full testing_start from Code to Run
# source("Code_to_Run.R", local = knitr::knit_global())
full_testing_start_date <- as.Date("2020-09-01", "%Y-%m-%d")

load(here("data", "table1_data.Rdata"))
#load(here("data", "table1_data_matched.Rdata"))
load(here("data", "table1_data_matched_age_week.Rdata"))

nice.num<-function(x){
  prettyNum(x, big.mark=",", nsmall = 0, digits=0,scientific = FALSE)}
nice.num2<-function(x){
  prettyNum(x, big.mark=",", nsmall = 2, digits=2,scientific = FALSE)}

```

```{r include=FALSE}
# variables for Table 1 
vars<-c(
        "age",
        "age_gr2",
        "gender",
        "trimester",
        "vaccination_status",
        "asthma",
        "autoimmune_disease",
        "copd",
        "dementia",
        "diabetes",
        "heart_disease",
        "cancer",
        "hypertension",
        "renal_impairment",
        "long_covid",
        "abdominal_pain",
        "allergy", 
        "altered_smell_taste",
        "anxiety" , 
        "blurred_vision" ,
        "chest_pain" ,
        "cognitive_dysfunction_brain_fog", 
        "cough" ,
        "depression" ,
        "dizziness" ,
        "dyspnea",
        "fatigue_malaise" ,
        "gastrointestinal_issues",
        "headache" ,
        "intermittent_fever",
        "joint_pain"  , 
        "memory_issues_v2" ,
        "menstrual_problems" ,
        "muscle_spams_pain",
        "neuralgia" , 
        "pins_sensation",  
        "sleep_disorder" ,
        "tachycardia" ,
        "tinnitus_hearing_problems", 
        "all_symp"
)

factor.vars<- c(
  "age_gr2",
        "gender",
        "trimester",
        "vaccination_status",
        "asthma",
        "autoimmune_disease",
        "copd",
        "dementia",
        "diabetes",
        "heart_disease",
        "cancer",
        "hypertension",
        "renal_impairment",
        "long_covid",
        "abdominal_pain",
        "allergy", 
        "altered_smell_taste",
        "anxiety" , 
        "blurred_vision" ,
        "chest_pain" ,
        "cognitive_dysfunction_brain_fog", 
        "cough" ,
        "depression" ,
        "dizziness" ,
        "dyspnea",
        "fatigue_malaise" ,
        "gastrointestinal_issues",
        "headache" ,
        "intermittent_fever",
        "joint_pain"  , 
        "memory_issues_v2" ,
        "menstrual_problems" ,
        "muscle_spams_pain",
        "neuralgia" , 
        "pins_sensation",  
        "sleep_disorder" ,
        "tachycardia" ,
        "tinnitus_hearing_problems", 
        "all_symp"
)

```

# Baseline characteristics 
```{r}
table1.data<-rbind(  
# those without missing values
       table1_data_new_infections   %>%
        mutate(group = "1 COVID-19 infection"),
       table1_data_confirmed_infection  %>%
        mutate(group = "2 Confirmed infection"),
       table1_data_first_infection %>%
        mutate(group = "3 First infection"),
       table1_data_reinfections  %>%
        mutate(group = "4 Reinfections"),
        table1_data_tested_negative_earliest  %>%
        mutate(group = "5 Tested negative, earliest"),
       table1_data_tested_negative_all %>%
        mutate(group = "6 Tested negative, all"),
       table1_data_pcr_negative_earliest  %>%
        mutate(group = "7 PCR negative, earliest"),
       table1_data_pcr_negative_all  %>%
        mutate(group = "8 PCR negative, all")) %>%
  filter(cohort_start_date >=full_testing_start_date)

summary.characteristics <-print(CreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = table1.data,
  strata=c("group"),
  test = F), 
  showAllLevels=F,
  smd=F,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:8) {
# tidy up 
  cur_column <- summary.characteristics[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics[, i] <- str_replace(string=summary.characteristics[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}
kable(summary.characteristics,
        col.names = c(
        "COVID-19 infection",
        "Confirmed infection",
        "First infection",
        "Reinfections",
        "Tested negative, earliest",
        "Tested negative, all ",
        "PCR negative, earliest ",
        "PCR negative, all "
                    ))%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

## New COVID-19 infection vs Tested negative (earliest event), after matching by week/year of testing and 5y age group
```{r include=FALSE}
data <- m.new_covid_tested_negative_earliest[[3]]

summary.characteristics_covid_negative <-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = data,
  strata=c("name"),
  test = F), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics_covid_negative[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics_covid_negative[, i] <- str_replace(string=summary.characteristics_covid_negative[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}

kable(summary.characteristics_covid_negative)%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

## Confirmed infections vs tested negative (earliest), after matching by week/year of testing and 5y age group
```{r}
data <- m.confirmed_infection_negative_earliest[[3]]

summary.characteristics.confirmed<-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = data,
  strata=c("name"),
  test = F), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics.confirmed[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics.confirmed[, i] <- str_replace(string=summary.characteristics.confirmed[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}
kable(summary.characteristics.confirmed
     )%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

## First infections vs reinfections, after matching by week/year of testing and 5y age group

```{r include=FALSE}

data <- m.first_infection_reinfections[[3]]

summary.characteristics.reinfections<-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = data,
  strata=c("name"),
  test = F), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics.reinfections[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics.reinfections[, i] <- str_replace(string=summary.characteristics.reinfections[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r echo=FALSE}

kable(summary.characteristics.reinfections 
     )%>%
  kable_styling(bootstrap_options = c("striped", "bordered")) #%>%
 # pack_rows("Baseline comorbidities", 40, 44)

```

```{r Old_function_matching, eval=FALSE, include=FALSE}
# matched_table_symptom_cohorts <- function(cohort_id1, cohort_id2, window, name1, name2){
#  results_list <- list()
#  
#   cohort1 <- cdm$er_long_covid_all_symptoms_cohorts %>%
#   filter(main_cohort_id == cohort_id1 & window_days == window) %>%
#   mutate(name = name1) %>%
#   mutate(cohort = "0") %>% 
#   distinct() %>%
#   collect()
# 
#   cohort2 <- cdm$er_long_covid_all_symptoms_cohorts %>%
#   filter(main_cohort_id == cohort_id2 & window_days == window) %>%
#    mutate(name = name2) %>%
#       mutate(cohort = "1") %>%
#   distinct() %>%
#   collect()
# 
#   data <- rbind(cohort1 , cohort2) %>% mutate(week_year = yearweek(as.Date(cohort_start_date)))
# # we match by week and year using exact method 
# m.week_year <- matchit(cohort ~ week_year, data = data, method = "exact")
# sum_matching <- summary(m.week_year)
# # we extract matched data
# m.data <- match.data(m.week_year) %>%
#   rename(ps_weights= weights) %>%
#     mutate(all_symp = as.factor(all_symp)) %>%
#   select(-subject_id, -window_days, - cohort_start_date , -week_year, -main_cohort_id, -cohort) 
# 
# s.data <- svydesign(ids = ~subclass, # clusters 
#           data = m.data, 
#           weights= ~ps_weights) # weights
# 
# summary <-  s.data%>% 
#    tbl_svysummary(
#        by = "cohort",
#   digits = all_categorical()~ c(0,1),
#   missing = "ifany",
#   missing_text= "Missing",
#   percent = "col"
#            ) %>% 
#   bold_labels()
# 
# results_list <- list(sum_matching, m.data, s.data,   summary)
# results_list
# }
# t.first.infections.reinfections <- matched_table_symptom_cohorts(cohort_id1= "70", name1= "First COVID-19 infection",
#                                                            cohort_id2 = "71", name2 = "Reinfections", 
#                                                            window="90")
# 
# 
# t.infections.tested.negative <- matched_table_symptom_cohorts(cohort_id1= "22", name1= "New COVID-19 infection",
#                                                            cohort_id2 = "30", name2 = "Tested negative, earliest", 
#                                                            window="90")

```

# 90 days Long Covid symptoms distribution
```{r}
vars <- c("abdominal_pain", "allergy", "altered_smell_taste",  "anxiety" , 
         "blurred_vision","chest_pain" ,"cognitive_dysfunction_brain_fog", "cough" ,"depression" ,
         "dizziness" ,"dyspnea" , "fatigue_malaise" , "gastrointestinal_issues","headache" ,"intermittent_fever" ,
         "joint_pain" ,  "memory_issues_v2" ,"menstrual_problems" , "muscle_spams_pain", "neuralgia" , "pins_sensation",  
         "sleep_disorder", "tachycardia" ,"tinnitus_hearing_problems")

factor.vars <- c("abdominal_pain", "allergy", "altered_smell_taste",  "anxiety" , 
         "blurred_vision","chest_pain" ,"cognitive_dysfunction_brain_fog", "cough" ,"depression" ,
         "dizziness" ,"dyspnea" , "fatigue_malaise" , "gastrointestinal_issues","headache" ,"intermittent_fever" ,
         "joint_pain" ,  "memory_issues_v2" ,"menstrual_problems" , "muscle_spams_pain", "neuralgia" , "pins_sensation",  
         "sleep_disorder", "tachycardia" ,"tinnitus_hearing_problems")

```

```{r include=FALSE}
# function to get data for the plot
get_prop <- function(data){
  filt <- c("1","2","3","4","5","6","7","8","9","10","11", "__all_symp__")
  names <- names(as_tibble(data,  col_labels = T)) 
  names <- str_extract(names, "\\*(.*)\\*")
  names <- str_sub(names, 3)
  names <- str_sub(names, end=-3)

  output <- as_tibble(data,  col_labels = F) %>%
   filter(!(label %in% filt)) %>%
    mutate(prop1 = str_extract(stat_1, "(?<=\\()(.*?)%")) %>%
    mutate(prop1= substring(prop1,1, nchar(prop1)-1)) %>%
  mutate(prop1 = as.numeric(prop1)) %>%
  select(-stat_1) %>%
     mutate(prop2 = str_extract(stat_2, "(?<=\\()(.*?)%")) %>%
  mutate(prop2= substring(prop2,1, nchar(prop2)-1)) %>%
  mutate(prop2 = as.numeric(prop2)) %>%
      select(-stat_2) %>%
  mutate(label= substring(label,1, nchar(label)-2)) %>%
  mutate(label= substring(label,3)) %>%
  mutate(label = str_replace_all(label, "_", " ")) %>%
  mutate(label = str_to_title(label)) %>%
  as.data.frame()

  names(output) <- names
  output
}

# plot function
function_matched_heatmap <- function(data){
 plot <-  data %>% 
  pivot_longer(!Characteristic, names_to = "Cohort", values_to = "Proportion") %>%
  mutate(Characteristic = fct_rev(factor(Characteristic))) %>%
  ggplot() +
  aes(x = 1, y = Characteristic, fill = Proportion) +
  facet_grid(~Cohort)+
  geom_tile(colour = "black") +
  scale_fill_gradient(high = "#C81D25", low = "#E4F1EB")+
  geom_text(aes(label= round(Proportion, 1))) +
  theme_minimal() +
  theme(axis.text.x=element_blank(),
        legend.position = "bottom")+
  xlab("")+
  ylab("Symptom")+
  labs(fill= "Proportion, in %")
 plot
}

```

## Matched New COVID-19 infection vs Tested negative (earliest record) cohorts
```{r echo=FALSE}

data <- m.new_covid_tested_negative_earliest[[2]] %>% filter(long_covid == 1) 
data <- data %>% 
  select(name, abdominal_pain, allergy, altered_smell_taste,  anxiety , 
         blurred_vision ,chest_pain ,cognitive_dysfunction_brain_fog, cough ,depression ,
         dizziness ,dyspnea , fatigue_malaise , gastrointestinal_issues,headache ,intermittent_fever ,
         joint_pain  ,  memory_issues_v2 ,menstrual_problems , muscle_spams_pain, neuralgia , pins_sensation,  
         sleep_disorder , tachycardia ,tinnitus_hearing_problems,  all_symp, ps_weights, matched_pair  )

s.data <- svydesign(ids = ~matched_pair,  
          data = data, 
          weights= ~ps_weights) # weights all 1

summary.characteristics_symptoms <-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = s.data,
  strata=c("name"),
  test = F), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics_symptoms[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
summary.characteristics_symptoms[, i] <- str_replace(string=summary.characteristics_symptoms[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}
kable(summary.characteristics_symptoms
     )%>%
  kable_styling(bootstrap_options = c("striped", "bordered")) #%>%
 # pack_rows("Baseline comorbidities", 40, 44)
```

## Heatmap

```{r eval=FALSE, include=FALSE}
results <-  get_prop(summary_new_covid_tested_negative) 
function_matched_heatmap(results)

```

## Matched Confirmed infection vs tested negative (earliest)
```{r}
data <- m.confirmed_infection_negative_earliest[[2]] %>% filter(long_covid == 1)
data <- data %>% 
  select(name, abdominal_pain, allergy, altered_smell_taste,  anxiety , 
         blurred_vision ,chest_pain ,cognitive_dysfunction_brain_fog, cough ,depression ,
         dizziness ,dyspnea , fatigue_malaise , gastrointestinal_issues,headache ,intermittent_fever ,
         joint_pain  ,  memory_issues_v2 ,menstrual_problems , muscle_spams_pain, neuralgia , pins_sensation,  
         sleep_disorder , tachycardia ,tinnitus_hearing_problems,  all_symp, ps_weights, matched_pair  )

s.data <- svydesign(ids = ~matched_pair,  
          data = data, 
          weights= ~ps_weights) # weights all 1

summary.characteristics_symptoms <-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = s.data,
  strata=c("name"),
  test = F), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics_symptoms[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
summary.characteristics_symptoms[, i] <- str_replace(string=summary.characteristics_symptoms[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}
kable(summary.characteristics_symptoms
     )%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))
```

```{r eval=FALSE, include=FALSE}
data <- m.confirmed_infection_negative_earliest[[2]] %>% filter(long_covid == 1)
# keep only variables of interest
data <- data %>% 
  select(name, abdominal_pain, allergy, altered_smell_taste,  anxiety , 
         blurred_vision ,chest_pain ,cognitive_dysfunction_brain_fog, cough ,depression ,
         dizziness ,dyspnea , fatigue_malaise , gastrointestinal_issues,headache ,intermittent_fever ,
         joint_pain  ,  memory_issues_v2 ,menstrual_problems , muscle_spams_pain, neuralgia , pins_sensation,  
         sleep_disorder , tachycardia ,tinnitus_hearing_problems,  all_symp, ps_weights, matched_pair  )

s.data <- svydesign(ids = ~matched_pair, # clusters 
          data = data, 
          weights= ~ps_weights) # weights

summary_confirmed_infection_tested_negative_earliest <-  s.data%>%
   tbl_svysummary(
       by = "name",
  digits = all_categorical()~ c(0,1),
  missing = "ifany",
  missing_text= "Missing",
  percent = "col"
           ) %>%
  bold_labels()

summary_confirmed_infection_tested_negative_earliest 
```

## Heatmap
```{r eval=FALSE, include=FALSE}
results <-  get_prop(summary_confirmed_infection_tested_negative_earliest ) 
function_matched_heatmap(results)
```

## Matched First COVID-19 infection vs Reinfection cohorts
```{r}
data <- m.first_infection_reinfections[[2]] %>% filter(long_covid == 1)
data <- data %>% 
  select(name, abdominal_pain, allergy, altered_smell_taste,  anxiety , 
         blurred_vision ,chest_pain ,cognitive_dysfunction_brain_fog, cough ,depression ,
         dizziness ,dyspnea , fatigue_malaise , gastrointestinal_issues,headache ,intermittent_fever ,
         joint_pain  ,  memory_issues_v2 ,menstrual_problems , muscle_spams_pain, neuralgia , pins_sensation,  
         sleep_disorder , tachycardia ,tinnitus_hearing_problems,  all_symp, ps_weights, matched_pair  )

s.data <- svydesign(ids = ~matched_pair,  
          data = data, 
          weights= ~ps_weights) # weights all 1

summary.characteristics_symptoms <-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = s.data,
  strata=c("name"),
  test = F), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics_symptoms[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
summary.characteristics_symptoms[, i] <- str_replace(string=summary.characteristics_symptoms[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}
kable(summary.characteristics_symptoms
     )%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))
```

## Heatmap
```{r eval=FALSE, include=FALSE}
results <-  get_prop(summary_first_infection_reinfections ) 
function_matched_heatmap(results)
```