### 2022-11-18 characterisation/Code/Comparing_symptoms_matched_nonbiased.Rmd
update

---
title: "Descriptive characteristics of Long COVID cohrots, matched"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning=FALSE, 
                      message=FALSE, 
                      comment = FALSE,
                      cache = FALSE,
                      cache.lazy=FALSE)
options(scipen=999)

```

```{r include=FALSE}
# packages
library(here)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(gtsummary)
library(tableone)
library(kableExtra)
library(ggplot2)
library(forcats)
library(tsibble)
library(survey)

# get data for table 1
# get full testing_start from Code to Run
# source("Code_to_Run.R", local = knitr::knit_global())
full_testing_start_date <- as.Date("2020-09-01", "%Y-%m-%d")

load(here("data", "table1_data.Rdata"))
load(here("data", "table1_data_matched.Rdata"))
#load(here("data", "table1_data_matched_age_week.Rdata"))

nice.num<-function(x){
  prettyNum(x, big.mark=",", nsmall = 0, digits=0,scientific = FALSE)}
nice.num2<-function(x){
  prettyNum(x, big.mark=",", nsmall = 2, digits=2,scientific = FALSE)}

```

```{r include=FALSE}
# variables for Table 1 
vars<-c(
        "age",
        "age_gr2",
        "gender",
        "trimester",
        "vaccination_status",
        "asthma",
        "autoimmune_disease",
        "copd",
        "dementia",
        "diabetes",
        "heart_disease",
        "cancer",
        "hypertension",
        "renal_impairment",
        "long_covid",
        "abdominal_pain",
        "allergy", 
        "altered_smell_taste",
        "anxiety" , 
        "blurred_vision" ,
        "chest_pain" ,
        "cognitive_dysfunction_brain_fog", 
        "cough" ,
        "depression" ,
        "dizziness" ,
        "dyspnea",
        "fatigue_malaise" ,
        "gastrointestinal_issues",
        "headache" ,
        "intermittent_fever",
        "joint_pain"  , 
        "memory_issues_v2" ,
        "menstrual_problems" ,
        "muscle_spams_pain",
        "neuralgia" , 
        "pins_sensation",  
        "sleep_disorder" ,
        "tachycardia" ,
        "tinnitus_hearing_problems", 
        "all_symp"
)

factor.vars<- c(
  "age_gr2",
        "gender",
        "trimester",
        "vaccination_status",
        "asthma",
        "autoimmune_disease",
        "copd",
        "dementia",
        "diabetes",
        "heart_disease",
        "cancer",
        "hypertension",
        "renal_impairment",
        "long_covid",
        "abdominal_pain",
        "allergy", 
        "altered_smell_taste",
        "anxiety" , 
        "blurred_vision" ,
        "chest_pain" ,
        "cognitive_dysfunction_brain_fog", 
        "cough" ,
        "depression" ,
        "dizziness" ,
        "dyspnea",
        "fatigue_malaise" ,
        "gastrointestinal_issues",
        "headache" ,
        "intermittent_fever",
        "joint_pain"  , 
        "memory_issues_v2" ,
        "menstrual_problems" ,
        "muscle_spams_pain",
        "neuralgia" , 
        "pins_sensation",  
        "sleep_disorder" ,
        "tachycardia" ,
        "tinnitus_hearing_problems", 
        "all_symp"
)

```

# Baseline characteristics 
```{r}
table1.data<-rbind(  
# those without missing values
       table1_data_new_infections   %>%
        mutate(group = "1 COVID-19 infection"),
       table1_data_confirmed_infection  %>%
        mutate(group = "2 Confirmed infection"),
       table1_data_first_infection %>%
        mutate(group = "3 First infection"),
       table1_data_reinfections  %>%
        mutate(group = "4 Reinfections"),
        table1_data_tested_negative_earliest  %>%
        mutate(group = "5 Tested negative, earliest"),
       table1_data_tested_negative_all %>%
        mutate(group = "6 Tested negative, all"),
       table1_data_pcr_negative_earliest  %>%
        mutate(group = "7 PCR negative, earliest"),
       table1_data_pcr_negative_all  %>%
        mutate(group = "8 PCR negative, all")) %>%
  filter(cohort_start_date >=full_testing_start_date) %>%
     mutate(abdominal_pain=ifelse(is.na(abdominal_pain),0,abdominal_pain)) %>% 
       mutate(allergy =ifelse(is.na(allergy ),0,allergy)) %>% 
       mutate(altered_smell_taste =ifelse(is.na(altered_smell_taste ),0,altered_smell_taste)) %>% 
       mutate(blurred_vision  =ifelse(is.na(blurred_vision  ),0,blurred_vision )) %>% 
       mutate( anxiety =ifelse(is.na(  anxiety),0, anxiety)) %>% 
       mutate(chest_pain =ifelse(is.na(chest_pain ),0,chest_pain)) %>% 
       mutate(cognitive_dysfunction_brain_fog=ifelse(is.na(cognitive_dysfunction_brain_fog),0,cognitive_dysfunction_brain_fog)) %>% 
       mutate( cough  =ifelse(is.na( cough  ),0, cough )) %>% 
       mutate(depression =ifelse(is.na(depression ),0,depression)) %>% 
       mutate(dizziness =ifelse(is.na(dizziness ),0,dizziness)) %>% 
       mutate(dyspnea =ifelse(is.na( dyspnea),0,dyspnea)) %>% 
       mutate(fatigue_malaise =ifelse(is.na(fatigue_malaise ),0,fatigue_malaise)) %>% 
       mutate(gastrointestinal_issues =ifelse(is.na(gastrointestinal_issues ),0,gastrointestinal_issues)) %>% 
       mutate(headache  =ifelse(is.na( headache ),0,headache )) %>% 
       mutate(intermittent_fever =ifelse(is.na(intermittent_fever ),0,intermittent_fever)) %>% 
       mutate( joint_pain =ifelse(is.na( joint_pain ),0, joint_pain)) %>% 
       mutate( memory_issues_v2  =ifelse(is.na( memory_issues_v2 ),0, memory_issues_v2 )) %>% 
       mutate(menstrual_problems =ifelse(is.na( menstrual_problems),0,menstrual_problems)) %>% 
       mutate( muscle_spams_pain =ifelse(is.na(  muscle_spams_pain),0, muscle_spams_pain)) %>% 
       mutate(neuralgia =ifelse(is.na(neuralgia),0,neuralgia)) %>% 
       mutate(pins_sensation =ifelse(is.na(pins_sensation ),0,pins_sensation)) %>% 
       mutate( sleep_disorder =ifelse(is.na(  sleep_disorder),0, sleep_disorder)) %>% 
       mutate(tachycardia =ifelse(is.na(tachycardia ),0,tachycardia)) %>% 
       mutate(tinnitus_hearing_problems =ifelse(is.na(tinnitus_hearing_problems ),0,tinnitus_hearing_problems)) %>% 
       mutate(all_symp =ifelse(is.na(all_symp ),0,all_symp))

summary.characteristics <-print(CreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = table1.data,
  strata=c("group"),
  test = F), 
  showAllLevels=F,
  smd=F,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:8) {
# tidy up 
  cur_column <- summary.characteristics[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics[, i] <- str_replace(string=summary.characteristics[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}
kable(summary.characteristics,
        col.names = c(
        "COVID-19 infection",
        "Confirmed infection",
        "First infection",
        "Reinfections",
        "Tested negative, earliest",
        "Tested negative, all ",
        "PCR negative, earliest ",
        "PCR negative, all "
                    ))%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

## New COVID-19 infection vs Tested negative (earliest event), after matching by week/year of testing and 5y age group
```{r include=FALSE}
data <- m.new_covid_tested_negative_earliest[[3]]

summary.characteristics_covid_negative <-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = data,
  strata=c("name"),
  test = T), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics_covid_negative[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics_covid_negative[, i] <- str_replace(string=summary.characteristics_covid_negative[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}

kable(summary.characteristics_covid_negative)%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

## Confirmed infections vs tested negative (earliest), after matching by week/year of testing and 5y age group
```{r}
data <- m.confirmed_infection_negative_earliest[[3]]

summary.characteristics.confirmed<-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = data,
  strata=c("name"),
  test = T), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics.confirmed[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics.confirmed[, i] <- str_replace(string=summary.characteristics.confirmed[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r}
kable(summary.characteristics.confirmed
     )%>%
  kable_styling(bootstrap_options = c("striped", "bordered"))

```

## First infections vs reinfections, after matching by week/year of testing and 5y age group

```{r include=FALSE}

data <- m.first_infection_reinfections[[3]]

summary.characteristics.reinfections<-print(svyCreateTableOne(
  vars =  vars,
  factorVars = factor.vars,
  includeNA=F,
  data = data,
  strata=c("name"),
  test = T), 
  showAllLevels=F,
  smd=T,
  nonnormal = vars, #all 
  noSpaces = TRUE,
  printToggle=FALSE)

for(i in 1:2) {
# tidy up 
  cur_column <- summary.characteristics.reinfections[, i]
  cur_column <- str_extract(cur_column, '[0-9.]+\\b') %>% 
                as.numeric() 
  cur_column <-nice.num(cur_column)
  # add back in
 summary.characteristics.reinfections[, i] <- str_replace(string=summary.characteristics.reinfections[, i],
                              pattern='[0-9.]+\\b', 
                              replacement=cur_column)    
}

```

```{r echo=FALSE}

kable(summary.characteristics.reinfections 
     )%>%
  kable_styling(bootstrap_options = c("striped", "bordered")) #%>%
 # pack_rows("Baseline comorbidities", 40, 44)

```